### 2018-09-28 FrascatiPlot/Paper/aps/main.tex
3737
Last minor fixes before submission

\input{table1}

For about sixty years, coalescence models have been used to describe the formation of composite objects \cite{Butler:1963, Kapusta:1980, Sato:1981ez, Nagle:1996vp, Scheibl:1998tk, Cho:2017dcy, Blum:2017qnn, Bazak:2018hgl, Zhao:2018lyf}.
Surprisingly, thermal-statistical models have been successful in describing also the production of light (anti-)(hyper-)nuclei across a wide range of energies in AA collisions \cite{Andronic:2010qu,Andronic:2017}. 
In this approach, particles are produced from a fireball in thermal and kinetic equilibrium with temperatures of the order of $T_{chem} \approx$ 156 MeV.
Particle abundances are fixed at chemical freeze-out, when inelastic collisions cease. Further elastic and pseudo-elastic collisions occur among the components of the expanding fireball, that can affect the spectral shapes and the measurable yields of short-lived (strongly decaying) hadronic resonances. Once the mean free path for elastic collisions is larger than the system size, the fireball freezes-out kinetically at $T_{kin} \approx$ 90 MeV~\cite{Abelev:2013vea}. 
In such a dense and hot environment, composite objects with binding energies that are small with respect to the temperature of the system, appear as ``fragile''. 
For instance, the binding energy of the deuteron is $B_{E, d}$ = 2.2 MeV $\ll T_{chem}, T_{kin}$.
The cross-section for pion-induced deuteron breakup is significantly larger than the typical (pseudo)-elastic cross-sections for the re-scattering of hadronic resonance decay products \cite{Garcilazo:1982yc, Bass:1998ca, Schukraft:2017nbn}. 
Similarly, the elastic cross-section driving deuteron spectra to kinetic equilibration in central heavy-ion collisions \cite{Acharya:2017dmc} is smaller than the breakup cross-section \cite{Garcilazo:1982yc, Bass:1998ca, Schukraft:2017nbn}.   
The deuterons produced at chemical freeze-out would be expected not to survive the hadronic phase, yet their production is measured to be consistent with statistical-thermal model predictions and a non-zero elliptic flow is observed \cite{Acharya:2017dmc}. 
Several solutions have been proposed to solve this ``light (anti-)nuclei puzzle'': (a.) a sudden freeze-out at the QGP-hadron phase boundary, (b.) the thermal production of these objects as compact quark bags \cite{Andronic:2017}, and (c.) the coincidence of coalescence and thermal production \cite{Scheibl:1998tk, HeinzTorino}. 
Data from rescattering of short-lived hadronic resonances suggest a long-lasting hadronic phase \cite{Abelev:2014uua}, thus strongly disfavouring hypothesis (a.). 
While hypothesis (b.) cannot presently be tested beyond the agreement of measured (anti-)nuclei yields with statistical-thermal model predictions, hypothesis (c.) is scrutinised in this letter.

To this purpose, we compare LHC data to models. 
For the first time, these data allow for the systematic study of the light (anti-)(hyper-)nuclei production as a function of the system- and object-size. 
A quantitative comparison of the two scenarios has been proposed in \cite{Mrowczynski:2016xqm}, resulting in the idea to study the production rates of nuclei with similar mass but very different internal structure, as \hefour~and ${}^{4}\mathrm{Li}$ \cite{Bazak:2018hgl}. However, as ${}^{4}\mathrm{Li}$ is not stable with respect to strong decay, its measurement is experimentally difficult and probably less constraining than the comparison with the hyper-nuclei proposed here.
We propose that \bA~is systematically measured in all collision systems by exploiting the large statistics sample foreseen at the LHC Run 3 \& 4, in order to discriminate the aforementioned scenarios.

\section{Coalescence approach} \label{sec:coalescence}

In the coalescence picture, nucleons produced in the collision coalesce into nuclei if they are close in space and have similar velocities \cite{Butler:1963,Kapusta:1980}. 
For a nucleus with mass number $A = Z + N$, the coalescence probability is quantified by the coalescence parameter $B_{A}$.
Considering that at LHC energies the number of produced protons and neutrons at midrapidity is expected to be equal,  $B_{A}$ is defined as

\begin{equation}
E_{A}\frac{\mathrm{d}^{3}N_{A}}{\mathrm{d}p_{A}^{3}}=B_{A}{\left(E_{\mathrm{p,n}}\frac{\mathrm{d}^{3}N_{\mathrm{p,n}}}{\mathrm{d}p_{\mathrm{p,n}}^{3}}\right)^{A}}\left\vert_{\vec{p}_{\mathrm{p}}=\vec{p}_{\mathrm{n}}=\frac{\vec{p}_{A}}{A}} \right..
\label{eq:BA}
\end{equation}
%
where $p_{\mathrm{p,n}}$ are the momenta of the proton and neutron and $E_{p,n}$ their energies.
%
The LHC is particularly suited for the production of anti-nuclei, since the number of baryons and anti-baryons is essentially equal at midrapidity \cite{Abbas:2013rua}. Consequently, the anti-particle to particle ratio for (hyper-)nuclei is consistent with unity \cite{ALICE:nucleipp2017, anielski-HQ14, Acharya:2017dmc, Adam:2015yta}.

In a simple coalescence approach, \bA~is expected to be independent of momentum and of the object size relative to the volume of particle emission (hereafter referred to as ``source volume'' or ``size'').
While this picture is found to be approximately valid in pp and \pPb~collisions \cite{ALICE:nucleipp2017, anielski-HQ14}, it breaks down in \PbPb~collisions, which exhibit a strong decrease of $B_{A}$ with centrality \cite{ALICE:deuteronppPbPb2015}. 
The elliptic flow of deuterons cannot be explained by simple coalescence \cite{Acharya:2017dmc}.

@@

The quantum-mechanical nature of the coalescence products is explicitly accounted for by means of an average correction factor, $\langle C_{A} \rangle$. For deuterons, $\langle C_{d} \rangle$ has been approximated as

\begin{equation}
\langle C_{d} \rangle \approx \frac{1}{\left[1+ \left(\frac{\dradius}{2R_{\perp}(m_{T})}\right)^{2}\right] \sqrt{1+ \left(\frac{\dradius}{2R_{\|}(m_{T})}\right)^{2}}}
\label{eq:Cd}
\end{equation}
%
where \dradius~is the size parameter, \rperp~and \rpar~are the lengths of homogeneity of the coalescence volume and $m_{T}$ is the transverse mass of the coalescing nucleons.
The nucleus size enters in the  calculation of \btwo~via $\langle C_{d} \rangle$, as well as the homogeneity volume $\rperp^{2}\rpar$, according to the relation~\cite{Scheibl:1998tk}
%
\begin{equation}
\btwo = \frac{3\pi^{3/2} \langle C_{d} \rangle}{2m_{T}\rperp^{2}(m_{T})\rpar(m_{T})} \;\; .
\label{eq:B2cd}
\end{equation}
%
The coalescence parameter decreases with increasing volume, as expected. 
$\langle C_{d} \rangle$ introduces a length scale defined by the deuteron size relative to the source size. 
If we assume that  $\rperp \approx \rpar \approx R$, Eqs. \ref{eq:Cd} and \ref{eq:B2cd} simplify to

\begin{equation}
\langle C_{d} \rangle \approx \left[1+ \left(\frac{\dradius}{2R(m_{T})}\right)^{2}\right]^{-3/2}
\label{eq:Cdapprox}
\end{equation}
%
and
%
\begin{equation}
\btwo = \frac{3\pi^{3/2} \langle C_{d} \rangle}{2m_{T}R^{3}(m_{T})}.
\label{eq:B2approx}
\end{equation}
%

Figure \ref{fig:radiusDependence} shows the source radius ($R$) dependence of $\langle C_{d} \rangle$ and \btwo, calculated assuming (a.) $\dradius = 0$, (b.) $\dradius = 0.3$ fm, (c.) the actual value $\dradius = 3.2$ fm~\cite{Mohr:2015ccw}, (d.) a larger, irrealistic $\dradius = 10$ fm. 
As shown in Fig. \ref{fig:radiusDependence}, $\langle C_{d} \rangle$ suppresses significantly the production of objects with radius larger than the source.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=.8\columnwidth]{../theory_coalescence_Cd_B2_vert.png}
\caption{{The quantum-mechanical correction factor $\langle C_{d} \rangle$ (left panel, see Eq.~\ref{eq:Cdapprox}) and the coalescence parameter \btwo~for deuteron (right panel, see Eq.~\ref{eq:B2approx}) as a function of the radius of the source $R$, calculated assuming a size parameter for the deuteron $\dradius = 0, 0.3, 3.2$ and $10$ fm. The inflection point of $\langle C_{d} \rangle$ corresponds to $R = \dradius/\sqrt{6}$ and is indicated in the left panel by the vertical dotted line for $\dradius = 3.2$ fm.}}
\label{fig:radiusDependence}
\end{center}
\end{figure}

Following the discussion in~\cite{Blum:2017qnn}, Eq.~\ref{eq:Cd} may be generalised as

\begin{equation}
\langle C_{A} \rangle = \prod_{i=1,2,3} \left(1 + \frac{r^2}{4R_{i}^2} \right)^{-\frac{1}{2}(A-1)} \;\; .
\label{eq:CA_general}
\end{equation}
%
Similarly, the coalescence parameter \bA~for a nucleus with mass number $A$ and spin $J_{A}$ is generalised by Eq. 6.2 of~\cite{Scheibl:1998tk}.
For the case of \hethree, the latter becomes Eq. 9 of~\cite{Blum:2017qnn}.

In summary, under the assumption $R_1\approx R_2 \approx R_3 \approx R$ as in \cite{Blum:2017qnn} (see next section) and by combining Eq. 6.2 in \cite{Scheibl:1998tk} and Eq.~\ref{eq:CA_general}, we obtain:

\begin{equation}
	\boxed {  B_{A} = {2J_{A} + 1 \over 2^{A}} {1 \over \sqrt{A}} {1 \over m_{T}^{A-1}} \Bigl({2\pi \over R^2 + ({r_A \over 2})^2 }\Bigr)^{3/2(A-1)} } \;.
\end{equation}

\noindent This general formula can be used to compare the predicted $B_{A}$ with data directly.

For small sources, as $R \rightarrow 0$, the coalescence probability is anti-proportional to the harmonic oscillator size parameter, and thus proportional to the depth of the attractive potential in the harmonic oscillator picture (and thus to the nucleus binding energy).
Quite naturally, the allowed momentum difference between the coalescing nucleons is larger for more attractive, i.e. deeper, potentials. 
For a large source, i.e. $R \gg r_A$, the coalescence probability is dominated by the classical phase-space separation, thus decreases for large distances in configuration space.

\subsection{Source volume}
\label{SecSourceVolume}
We identify the source volume as the effective sub-volume of the whole system that is governed by the homogeneity length of the interacting nucleons, as in \cite{Scheibl:1998tk}, and experimentally accessible with Hanbury-Brown-Twiss (HBT) interferometry. 
Experimental results are typically obtained following the Bertsch-Pratt (BP) parameterisation ($R_{out}, R_{side}, R_{long}$), while the coalescence model described in Section \ref{sec:coalescence} expresses the volume in terms of the Yano-Koonin-Podgoretskii (YKP) parameterisation. 
We identify $\rperp = R_{side}$, $\rpar = R_{long}$ and then take $R = (\rperp^{2}\rpar)^{1/3} \approx (R_{side}^{2}R_{long})^{1/3}$.

Experimentally, $R$ can be controlled by selecting different collision geometries, i.e. different centrality classes \cite{Abelev:2013qoq}. 
The HBT radii scale with the cubic root of the average charged-particle multiplicity density \avdNdeta$^{1/3}$ \cite{Adam:2015vna}, and depend on the pair average transverse momentum $\langle k_{\mathrm{T}}\rangle$ \cite{Aamodt:2011mr}. We make the simplifying assumption that the scaling with \avdNdeta$^{1/3}$ holds across collision systems, which is approximately fulfilled in data \cite{Adam:2015pya}. 
In contrast to \cite{Blum:2017qnn}, we do not explicitly use the measured HBT radii in our study, but we derive$R$ from the measured \avdNdeta~according to the following:

\begin{equation}
R = a \avdNdeta^{1/3} + b \;\;.
\label{eq:radiusParameterisation}
\end{equation}

The coefficients, $a = 0.339$ and $b = 0.128$ (in units of fm), have been determined by fitting linearly the ALICE data, as shown in Fig. \ref{fig:radiiparam}. 
The values we obtain by interpolating the geometric mean of the measured radii are consistent with the radii from kaon femtoscopy for $m_{\mathrm{T}} \approx 1$ \GeVc~in low-multiplicity pp collisions \cite{Abelev:2012sq} and the radii from pion femtoscopy in high-multiplicity \PbPb~collisions at the highest available $k_{\mathrm{T}} \approx 0.9$ \GeVc~\cite{Adam:2015vna}. 
The highest $k_{\mathrm{T}}$ was chosen as it is closest in $m_{\mathrm{T}}$ to the lowest transverse momentum per nucleon (\pt/$A \approx 0.8$ \GeVc) accessible by ALICE for the measurement of nuclei production. 
Ideally, one would use the proton femtoscopic radii, but given the unavailability of these measurements in every collision system and centrality, we assume that $m_{\mathrm{T}}$-scaling holds\cite{Adam:2015vja}.

@@

\section{Statistical-thermal approach and blast-wave}\label{sec:thermal}

In the statistical-thermal approach \cite{Petran:2013dva,Wheaton:2004qb,Andronic:2005yp}, the yields (d$N$/d$y$) of light anti- and hyper-nuclei are very sensitive to $T_{chem}$ due to their large mass and approximately scale as d$N$/d$y \propto \exp(-m/T_{chem})$. 
At the LHC, the chemical potentials which ensure the conservation of baryon number, strangeness, and electric charge are negligible.
In contrast to coalescence, the statistical-thermal models provide only \pt-integrated yields. 
We use a blast-wave~\cite{Schnedermann:1993ws} parameterization to model the \pt-dependence, with parameters obtained from the simultaneous fit to pion, kaon and proton spectra measured in Pb--Pb collisions by ALICE~\cite{Abelev:2013vea}. 
The normalisation is fixed using the \pt-integrated deuteron-to-pion ratio and \hethree-to-pion ratio predicted by the GSI-Heidelberg model with $T_{chem} = 156$ MeV, multiplied by the measured pion yield~\cite{Abelev:2013vea}. 
This choice, as opposed to using the ratio to proton, is motivated by the fact that the measured proton yield is seen to be slightly overestimated by the thermal model~\cite{Abelev:2012wca}.
For hyper-triton, the normalisation is extracted from the statistical-thermal model prediction of the strangeness population factor $S_{3} = { _{\Lambda}^{3}\mathrm{H} / ^{3}{\mathrm{He}} \over \Lambda/p}$ multiplied by the measured $\Lambda/p$ ratio~\cite{Abelev:2013vea,Abelev:2013xaa} and \hethree~yield~\cite{ALICE:deuteronppPbPb2015}. 
With the resulting spectra, we calculate \bA~for a given \pt/$A$ and compare it with coalescence expectations. 
We use the corresponding \avdNdeta~in each class to estimate the system radius based on the parameterisation discussed in Sec.~\ref{SecSourceVolume}. 
In contrast to the coalescence approach, the object size does not enter in the formulation of the blast-wave model.
The thermal model on the other hand, implements eigenvolume corrections by fixing the object radius as an external parameter.
We refer to the literature for the extensive discussions on the validity of the eigenvolume correction for light (anti-)(hyper-)nuclei~\cite{Vovchenko:2016mwg} and the relation with the possible production as compact quark bags~\cite{Andronic:2017}.

@@

\section{Comparison with data}

>>>>

For about sixty years, coalescence models have been used to describe the formation of composite objects \cite{Butler:1963, Kapusta:1980, Sato:1981ez, Nagle:1996vp, Scheibl:1998tk, Cho:2017dcy, Blum:2017qnn, Bazak:2018hgl, Zhao:2018lyf}.
Surprisingly, thermal-statistical models have been successful in describing also the production of light \mbox{(anti-)}(hyper-)nuclei across a wide range of energies in AA collisions \cite{Andronic:2010qu,Andronic:2017}. 
In this approach, particles are produced from a fireball in thermal and kinetic equilibrium with temperatures of the order of $T_{chem} \approx$ 156 MeV.
Particle abundances are fixed at chemical freeze-out, when inelastic collisions cease. Further elastic and pseudo-elastic collisions occur among the components of the expanding fireball, that can affect the spectral shapes and the measurable yields of short-lived (strongly decaying) hadronic resonances. Once the mean free path for elastic collisions is larger than the system size, the fireball freezes-out kinetically at $T_{kin} \approx$ 90 MeV~\cite{Abelev:2013vea}. 
In such a dense and hot environment, composite objects with binding energies that are small with respect to the temperature of the system, appear as ``fragile''. 
For instance, the binding energy of the deuteron is $B_{E, d}$ = 2.2~MeV~$\ll~T_{chem},~T_{kin}$.
The cross-section for pion-induced deuteron breakup is significantly larger than the typical (pseudo)-elastic cross-sections for the re-scattering of hadronic resonance decay products \cite{Garcilazo:1982yc, Bass:1998ca, Schukraft:2017nbn}. 
Similarly, the elastic cross-section driving deuteron spectra to kinetic equilibration in central heavy-ion collisions \cite{Acharya:2017dmc} is smaller than the breakup cross-section \cite{Garcilazo:1982yc, Bass:1998ca, Schukraft:2017nbn}.   
The deuterons produced at chemical freeze-out would be expected not to survive the hadronic phase, yet their production is measured to be consistent with statistical-thermal model predictions and a non-zero elliptic flow is observed \cite{Acharya:2017dmc}. 
Several solutions have been proposed to solve this ``light (anti-)nuclei puzzle'': (a.) a sudden freeze-out at the QGP-hadron phase boundary, (b.) the thermal production of these objects as compact quark bags \cite{Andronic:2017}, and (c.) the coincidence of coalescence and thermal production \cite{Scheibl:1998tk, HeinzTorino}. 
Data from rescattering of short-lived hadronic resonances suggest a long-lasting hadronic phase \cite{Abelev:2014uua}, thus strongly disfavouring hypothesis (a.). 
While hypothesis (b.) cannot presently be tested beyond the agreement of measured (anti-)nuclei yields with statistical-thermal model predictions, hypothesis (c.) is scrutinised in this letter.
%
\begin{table*}[htb]
\centering
\begin{ruledtabular}
\begin{tabular}{cccccccc}\\[-2ex]
Mass number & Nucleus &  Composition  & Binding energy (MeV)   &  Spin & \rmsradius$^{meas}$~(fm) &  $r_{A}$ (fm) & Refs. 
\\[0.5ex] \hline \\[-2ex]
      A = 2  & d                                    & pn                   &   2.224575 (9)     &     1   & 2.1413 $\pm$ 0.0025      &  3.2    &   \cite{VanDerLeun:1982bhg,Mohr:2015ccw}     \\[0.5ex]  \hline \\[-2ex]
      A = 3  & \tritium 	                  & pnn                       &    8.4817986 (20) & 1/2   &  1.755  $\pm$ 0.086        &  2.15   &   \cite{Purcell:2015gtm}           \\
                & \hethree                         & ppn                 &   7.7180428  (23) & 1/2   & 1.959 $\pm$  0.030         &   2.48  &   \cite{Purcell:2015gtm} \\
                & \hthreelambda               & p$\Lambda$n &    0.13 $\pm$ 0.05 & 1/2  &  4.9 --  10.0                     &  6.8 -- 14.1 & \cite{Davis:2005mb,Nemura:1999qp} \\[0.5ex]                                    
\end{tabular}
\end{ruledtabular}
\caption{Properties of nuclei and hyper-nuclei with mass number $A \leq 3$. The nucleus size is given in terms of the (charge) rms radius of the wave-function, \rmsradius. The size parameter of the wave-function of the harmonic oscillator potential,  $r_{A}$, is chosen such that the measured/expected rms,  \rmsradius$^{meas}$~(fm), is approximately reproduced. The proton rms charge radius $\lambda_{p} = 0.879(8)$ fm \cite{bernauer10} is subtracted from $\rmsradius^{meas}$  according to $\rmsradius = \sqrt{(\rmsradius^{meas})^2 - {\lambda_{p}^2}}$ to account for the finite extension of the constituents. We assume $\lambda_{\Lambda}\approx \lambda_{n}\approx \lambda_{p}$.}
\label{tab:nucleusradii}
\end{table*}%

\indent To this purpose, we compare LHC data to models. 
For the first time, these data allow for the systematic study of the light (anti-)(hyper-)nuclei production as a function of the system- and object-size. 
A quantitative comparison of the two scenarios has been proposed in \cite{Mrowczynski:2016xqm}, resulting in the idea  to study the production rates of nuclei with similar mass but very different internal structure, as \hefour~and ${}^{4}\mathrm{Li}$ \cite{Bazak:2018hgl}. However, as ${}^{4}\mathrm{Li}$ is not stable with respect to strong decay, its measurement is experimentally difficult and probably less constraining than the comparison with the hyper-nuclei proposed here.
We propose that \bA~is systematically measured in all collision systems by exploiting the large statistics sample foreseen at the High-Luminosity LHC, in order to discriminate the aforementioned scenarios.

\section{Coalescence approach} \label{sec:coalescence}
In the coalescence picture, nucleons produced in the collision coalesce into nuclei if they are close in space and have similar velocities \cite{Butler:1963,Kapusta:1980}. 
For a nucleus with mass number $A = Z + N$, the coalescence probability is quantified by the coalescence parameter $B_{A}$.
Considering that at LHC energies the number of produced protons and neutrons at midrapidity is expected to be equal,  $B_{A}$ is defined as

\begin{equation}
E_{A}\frac{\mathrm{d}^{3}N_{A}}{\mathrm{d}p_{A}^{3}}=B_{A}{\left(E_{\mathrm{p,n}}\frac{\mathrm{d}^{3}N_{\mathrm{p,n}}}{\mathrm{d}p_{\mathrm{p,n}}^{3}}\right)^{A}}\left\vert_{\vec{p}_{\mathrm{p}}=\vec{p}_{\mathrm{n}}=\frac{\vec{p}_{A}}{A}} \right.,
\label{eq:BA}
\end{equation}
%
where $p_{\mathrm{p,n}}$ are the momenta of the proton and neutron and $E_{p,n}$ their energies.
%
The LHC is particularly suited for the production of anti-nuclei, since the number of baryons and anti-baryons is essentially equal at midrapidity \cite{Abbas:2013rua}. Consequently, the anti-particle to particle ratio for (hyper-)nuclei is consistent with unity \cite{ALICE:nucleipp2017, anielski-HQ14, Acharya:2017dmc, Adam:2015yta}.
In a simple coalescence approach, \bA~is expected to be independent of momentum and of the object size relative to the volume of particle emission (hereafter referred to as ``source volume'' or ``size'').
While this picture is found to be approximately valid in pp and \pPb~collisions \cite{ALICE:nucleipp2017, anielski-HQ14}, it breaks down in \PbPb~collisions, which exhibit a strong decrease of $B_{A}$ with centrality \cite{ALICE:deuteronppPbPb2015}. 
The elliptic flow of deuterons cannot be explained by simple coalescence \cite{Acharya:2017dmc}.

@@

The quantum-mechanical nature of the coalescence products is explicitly accounted for by means of an average correction factor, $\langle C_{A} \rangle$. For deuterons, $\langle C_{d} \rangle$ has been approximated as 
%
\begin{equation}
\langle C_{d} \rangle \approx \frac{1}{\left[1+ \left(\frac{\dradius}{2R_{\perp}(m_{T})}\right)^{2}\right] \sqrt{1+ \left(\frac{\dradius}{2R_{\|}(m_{T})}\right)^{2}}}
\label{eq:Cd}
\end{equation}
%
where \dradius~is the size parameter, \rperp~and \rpar~are the lengths of homogeneity of the coalescence volume and $m_{T}$ is the transverse mass of the coalescing nucleons.
The nucleus size enters in the  calculation of \btwo~via $\langle C_{d} \rangle$, as well as the homogeneity volume $\rperp^{2}\rpar$, according to the relation~\cite{Scheibl:1998tk}
%
\begin{equation}
\btwo = \frac{3\pi^{3/2} \langle C_{d} \rangle}{2m_{T}\rperp^{2}(m_{T})\rpar(m_{T})} \;\; .
\label{eq:B2cd}
\end{equation}
%
The coalescence parameter decreases with increasing volume, as expected. 
$\langle C_{d} \rangle$ introduces a length scale defined by the deuteron size relative to the source size. 
If we assume that  $\rperp \approx \rpar \approx R$, Eqs. \ref{eq:Cd} and \ref{eq:B2cd} simplify to 
%
\begin{equation}
\langle C_{d} \rangle \approx \left[1+ \left(\frac{\dradius}{2R(m_{T})}\right)^{2}\right]^{-3/2}
\label{eq:Cdapprox}
\end{equation}
%
and
%
\begin{equation}
\btwo = \frac{3\pi^{3/2} \langle C_{d} \rangle}{2m_{T}R^{3}(m_{T})}.
\label{eq:B2approx}
\end{equation}
%
Figure \ref{fig:radiusDependence} shows the source radius ($R$) dependence of $\langle C_{d} \rangle$ and \btwo, calculated assuming (a.) $\dradius = 0$, (b.) $\dradius = 0.3$ fm, (c.) the actual value $\dradius = 3.2$ fm~\cite{Mohr:2015ccw}, (d.) a larger, irrealistic $\dradius = 10$ fm. 
As shown in Fig. \ref{fig:radiusDependence}, $\langle C_{d} \rangle$ suppresses significantly the production of objects with radius larger than the source.

\begin{figure}[hb]
\begin{center}
\includegraphics[width=\columnwidth]{../theory_coalescence_Cd_B2_vert.png}
\caption{{The quantum-mechanical correction factor $\langle C_{d} \rangle$ (top, see Eq.~\ref{eq:Cdapprox}) and the coalescence parameter \btwo~for deuteron (bottom, see Eq.~\ref{eq:B2approx}) as a function of the radius of the source $R$, calculated assuming a size parameter for the deuteron $\dradius = 0, 0.3, 3.2$ and $10$ fm. The inflection point of $\langle C_{d} \rangle$ corresponds to $R = \dradius/\sqrt{6}$ and is indicated in the left panel by the vertical dotted line for $\dradius = 3.2$ fm.}}
\label{fig:radiusDependence}
\end{center}
\end{figure}

Following the discussion in~\cite{Blum:2017qnn}, Eq.~\ref{eq:Cd} may be generalised as 
%
\begin{equation}
\langle C_{A} \rangle = \prod_{i=1,2,3} \left(1 + \frac{r^2}{4R_{i}^2} \right)^{-\frac{1}{2}(A-1)} \;\; .
\label{eq:CA_general}
\end{equation}
%
Similarly, the coalescence parameter \bA~for a nucleus with mass number $A$ and spin $J_{A}$ is generalised by Eq. 6.2 of~\cite{Scheibl:1998tk}.
For the case of \hethree, the latter becomes Eq. 9 of~\cite{Blum:2017qnn}.

In summary, under the assumption $R_1\approx R_2 \approx R_3 \approx R$ as in \cite{Blum:2017qnn} (see next section) and by combining Eq. 6.2 in \cite{Scheibl:1998tk} and our Eq.~\ref{eq:CA_general}, we obtain:
%
\begin{equation}
	\boxed {  B_{A} = {2J_{A} + 1 \over 2^{A}} {1 \over \sqrt{A}} {1 \over m_{T}^{A-1}} \Bigl({2\pi \over R^2 + ({r_A \over 2})^2 }\Bigr)^{3/2(A-1)} } \;.
\end{equation}
%
This general formula can be used to compare the predicted $B_{A}$ with data directly.

For small sources, as $R \rightarrow 0$, the coalescence probability is anti-proportional to the harmonic oscillator size parameter, and thus proportional to the depth of the attractive potential in the harmonic oscillator picture (and thus to the nucleus binding energy).
Quite naturally, the allowed momentum difference between the coalescing nucleons is larger for more attractive, i.e. deeper, potentials. 
For a large source, i.e. $R \gg r_A$, the coalescence probability is dominated by the classical phase-space separation, thus decreases for large distances in configuration space. 
%

\section{Source volume} \label{SecSourceVolume}
We identify the source volume as the effective sub-volume of the whole system that is governed by the homogeneity length of the interacting nucleons, as in \cite{Scheibl:1998tk}, and experimentally accessible with Hanbury-Brown-Twiss (HBT) interferometry. 
Experimental results are typically obtained following the Bertsch-Pratt (BP) parameterisation ($R_{out}, R_{side}, R_{long}$), while the coalescence model described in Section \ref{sec:coalescence} expresses the volume in terms of the Yano-Koonin-Podgoretskii (YKP) parameterisation. 
We identify $\rperp = R_{side}$, $\rpar = R_{long}$ and then take $R = (\rperp^{2}\rpar)^{1/3} \approx (R_{side}^{2}R_{long})^{1/3}$.

Experimentally, $R$ can be controlled by selecting different collision geometries, i.e. different centrality classes \cite{Abelev:2013qoq}. 
The HBT radii scale with the cubic root of the average charged-particle multiplicity density \avdNdeta$^{1/3}$ \cite{Adam:2015vna}, and depend on the pair average transverse momentum $\langle k_{\mathrm{T}}\rangle$ \cite{Aamodt:2011mr}. We make the simplifying assumption that the scaling with \avdNdeta$^{1/3}$ holds across collision systems, which is approximately fulfilled in data \cite{Adam:2015pya}. 
In contrast to \cite{Blum:2017qnn}, we do not explicitly use the measured HBT radii in our study, but we derive $R$ from the measured \avdNdeta~according to the following:
%
\begin{equation}
R = a \avdNdeta^{1/3} + b \;\;.
\label{eq:radiusParameterisation}
\end{equation}
%
The coefficients, $a = 0.339$ and $b = 0.128$ (in units of fm), have been determined by fitting linearly the ALICE data, as shown in Fig. \ref{fig:radiiparam}. 
The values we obtain by interpolating the geometric mean of the measured radii are consistent with the radii from kaon femtoscopy for $m_{\mathrm{T}} \approx 1$ \GeVc~in low-multiplicity pp collisions \cite{Abelev:2012sq} and the radii from pion femtoscopy in high-multiplicity \PbPb~collisions at the highest available $k_{\mathrm{T}} \approx 0.9$ \GeVc~\cite{Adam:2015vna}. 
The highest $k_{\mathrm{T}}$ was chosen as it is closest in $m_{\mathrm{T}}$ to the lowest transverse momentum per nucleon (\pt/$A \approx 0.8$ \GeVc) accessible by ALICE for the measurement of nuclei production. 
Ideally, one would use the proton femtoscopic radii, but given the unavailability of these measurements in every collision system and centrality, we assume that $m_{\mathrm{T}}$-scaling holds~\cite{Adam:2015vja}.

@@

\section{Statistical-thermal approach and blast-wave}\label{sec:thermal}
In the statistical-thermal approach \cite{Petran:2013dva,Wheaton:2004qb,Andronic:2005yp}, the yields (d$N$/d$y$) of light anti- and hyper-nuclei are very sensitive to $T_{chem}$ due to their large mass and approximately scale as d$N$/d$y \propto \exp(-m/T_{chem})$. 
At the LHC, the chemical potentials which ensure the conservation of baryon number, strangeness, and electric charge are negligible.
In contrast to coalescence, the statistical-thermal models provide only \pt-integrated yields. 
We use a blast-wave~\cite{Schnedermann:1993ws} parameterization to model the \pt-dependence, with parameters obtained from the simultaneous fit to pion, kaon and proton spectra measured in Pb--Pb collisions by ALICE~\cite{Abelev:2013vea}. 
The normalisation is fixed using the \pt-integrated deuteron-to-pion ratio and \hethree-to-pion ratio predicted by the GSI-Heidelberg model with $T_{chem} = 156$ MeV, multiplied by the measured pion yield~\cite{Abelev:2013vea}. 
This choice, as opposed to using the ratio to proton, is motivated by the fact that the measured proton yield is seen to be slightly overestimated by the thermal model~\cite{Abelev:2012wca}.
For hyper-triton, the normalisation is extracted from the statistical-thermal model prediction of the strangeness population factor $S_{3} = { _{\Lambda}^{3}\mathrm{H} / ^{3}{\mathrm{He}} \over \Lambda/p}$ multiplied by the measured $\Lambda/p$ ratio~\cite{Abelev:2013vea,Abelev:2013xaa} and \hethree~yield~\cite{ALICE:deuteronppPbPb2015}. 
With the resulting spectra, we calculate \bA~for a given \pt/$A$ and compare it with coalescence expectations. 
We use the corresponding \avdNdeta~in each class to estimate the system radius based on the parameterisation discussed in Sec.~\ref{SecSourceVolume}. 
In contrast to the coalescence approach, the object size does not enter in the formulation of the blast-wave model.
The thermal model on the other hand, implements eigenvolume corrections by fixing the object radius as an external parameter.
We refer to the literature for the extensive discussions on the validity of the eigenvolume correction for light \mbox{(anti-)(hyper-)}nuclei~\cite{Vovchenko:2016mwg} and the relation with the possible production as compact quark bags~\cite{Andronic:2017}.

@@

 \begin{figure}[ht]
	\begin{center}
		\includegraphics[width=.95\columnwidth]{../compareThermalAndCoalescence.png}
		\caption{Comparison of the coalescence parameters measured by ALICE (solid markers) for deuterons (upper panel), \hethree\ (middle panel) and \hthreelambda\ (lower panel) in pp \cite{ALICE:nucleipp2017} and Pb--Pb~\cite{ALICE:deuteronppPbPb2015, Adam:2015yta} collisions with the  thermal+blast-wave model expectations (dotted line) and the coalescence predictions (solid lines). Parameterisation B has been used to map \avdNdeta~into the radius $R$ of the source. The dashed line in the lower right panel corresponds to the coalescence prediction for the \hthreelambda\ with a larger radius.}
		\label{fig:CompareThermalAndCoalescence}
	\end{center}
\end{figure}

\section{Comparison with data}
In Fig.~\ref{fig:CompareThermalAndCoalescence}, the available data for (anti-)deuterons, \mbox{(anti-)}\hethree\ and (anti-)\hthreelambda~\cite{Adam:2015yta} are compared to coalescence and to the thermal model+blast wave predictions described in Sec.~\ref{sec:thermal}.
For the latter and for data, the radius parameterisation B is used, see Sec.~\ref{sec:radiiParamet}. 
For deuterons, both approaches lead to similar predictions and give a reasonable description of the experimental data for $R \gtrsim 1.6$ fm. 
For \hethree~the two models exhibit a qualitatively similar trend as a function of $R$, but they differ by a factor of about 1.5 to 2. The limited amount of data currently available is consistent with both models within 2$\sigma$ to 3$\sigma$, where $\sigma$ is the total uncertainty on the data. 
Both approaches show large differences (a factor 5 to 6 for central Pb--Pb collisions and a factor larger than 50 for small radii, $R < 2$ fm) for the \hthreelambda\ caused by the significantly larger size of \hthreelambda\ with respect to \hethree. 
The only data point available so far in Pb--Pb collisions is in agreement with the thermal model but not with coalescence. 
The difference between data and the coalescence model is about 6$\sigma$,
albeit the validity of our assumptions (\textit{in primis} the usage of a gaussian wave-function).
In \cite{Zhang:2018euf} it is argued that the difference between data and the coalescence model might be explained by a later formation through coalescence of $\Lambda$s and deuterons. 
A possible difference attributable to the presence of excited states with $J=3/2$ of \hthreelambda, which would significantly enhance the phase space for its production, is not considered here as there is no evidence for its existence~\cite{Mart:1996ay}.
